---
title: 'Statistical Inference Course Project: Part 2'
output: 
  pdf_document: 
    latex_engine: xelatex
---

*by Maxim Podkolzine*

\vspace*{1\baselineskip}
**Problem:** Analyze the `ToothGrowth` data in the R datasets package.

\vspace*{1\baselineskip}
**Q1-Q2:** First of all let's load the data and provide a basic summary.

```{r, echo=FALSE}
library(datasets)
library(ggplot2)
```

```{r}
data(ToothGrowth)
head(ToothGrowth)
```

The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

```{r}
str(ToothGrowth)
```

A data frame contains 60 observations of 3 variables: `len`	is a tooth length (numeric), `supp`	 is a	supplement type (a factor, VC or OJ), `dose`	is a dose in milligrams (numeric, 0.5, 1 or 2).

```{r}
summary(ToothGrowth)
```

A visual presentation might be helpful.

```{r}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, color=supp)) +
  geom_point(size=3, alpha=0.6) +
  facet_grid(. ~ supp) +
  xlab("Dose") +
  ylab("Tooth length") +
  guides(color=guide_legend(title="Supplement type"))
```

\vspace*{1\baselineskip}
**Q3:** Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

As can be seen in the plot, there is a clear positive correlation between the tooth length and the dose levels of Vitamin C, for both delivery methods. Let's test that formally, but start with a simple regression.

```{r}
fit <- lm(len ~ dose + supp, data=ToothGrowth)
summary(fit)
```

`R-squared` is equal to 70%, which means most of the variance in tooth length is explained by the dose and supplement.

<br>
Now let's t-test that the tooth length depends on the dosage.

```{r}
dose.0.5 = ToothGrowth$len[ToothGrowth$dose == 0.5]
dose.1 = ToothGrowth$len[ToothGrowth$dose == 1]
dose.2 = ToothGrowth$len[ToothGrowth$dose == 2]
```

```{r}
t.test(dose.1, dose.0.5, paired=FALSE, var.equal=TRUE)
t.test(dose.2, dose.1, paired=FALSE, var.equal=TRUE)
```

95% confidence intervals for both tests do not contain 0. So we reject the null hypothesis (which is the difference in length is 0), concluding that the dosage effect is significant.

\vspace*{1\baselineskip}
Finally let's t-test that the tooth length depends on the delivery method.

```{r}
oj.group = ToothGrowth$len[ToothGrowth$supp=="OJ"]
vc.group = ToothGrowth$len[ToothGrowth$supp=="VC"]
```

```{r}
t.test(oj.group, vc.group, paired=FALSE, var.equal=FALSE)
```

95% confidence interval does contain 0, so we cannot reject the null hypothesis (which is again the difference in length is 0).
But we can with 90% interval, since the p-value is 6%.

\vspace*{1\baselineskip}
**Q4:** Conclusions.

- Dosage has positive impact on tooth length: the length increases with higher dosage of Vitamin C.
- Orange juice is more eﬀective than ascorbic acid, but the influence is not that significant, mostly because the difference with 2mg dosage is very small.